---
title: "Sandbox Rei"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(forcats)
#library(plyr)
library(plotly)
```


```{r}
park_colors <- c("darkgreen", "cornflowerblue", "chocolate1", "blueviolet")
temp_colors <- c("red", "orange", "yellow", "blue", "cyan")

temp_list_names <- c("xhot_days", "hot_days", "normal_days", "cool_days", "xcool_days")
days_of_week <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
```


```{r}
dwarfs_train <- read_csv("../data/7_dwarfs_train.csv")
alien_saucers <- read_csv("../data/alien_saucers.csv")
dinosaur <- read_csv("../data/dinosaur.csv")
expedition_everest <- read_csv("../data/expedition_everest.csv")
flight_of_passage <- read_csv("../data/flight_of_passage.csv")
kilimanjaro_safaris <- read_csv("../data/kilimanjaro_safaris.csv")
navi_river <- read_csv("../data/navi_river.csv")
pirates_of_caribbean <- read_csv("../data/pirates_of_caribbean.csv")
rock_n_rollercoaster <- read_csv("../data/rock_n_rollercoaster.csv")
slinky_dog <- read_csv("../data/slinky_dog.csv")
soarin <- read_csv("../data/soarin.csv")
spaceship_earth <- read_csv("../data/spaceship_earth.csv")
splash_mountain <- read_csv("../data/splash_mountain.csv")
toy_story_mania <- read_csv("../data/toy_story_mania.csv")

head(dwarfs_train)
```


```{r}
rides <- list("dwarfs_train" = dwarfs_train, 
              "alien_saucers" = alien_saucers,
              "dinosaur" = dinosaur,
              "expedition_everest" = expedition_everest,
              "flight_of_passage" = flight_of_passage,
              "kilimanjaro_safaris" = kilimanjaro_safaris,
              "navi_river" = navi_river,
              "pirates_of_caribbean" = pirates_of_caribbean,
              "rock_n_rollercoaster" = rock_n_rollercoaster,
              "slinky_dog" = slinky_dog,
              "soarin" = soarin,
              "spaceship_earth" = spaceship_earth,
              "splash_mountain" = splash_mountain,
              "toy_story_mania" = toy_story_mania)

head(rides[1])
```

```{r}
for (i in 1:14) {
  rides[[i]] <- rides[[i]] %>%
      mutate(time = format(ymd_hms(datetime), "%H:%M:%S"))
  rides[[i]] <- rides[[i]] %>%
    mutate(ride_name = as.factor(names(rides[i])))
    rides[[i]] <- rides[[i]] %>%
    filter(SPOSTMIN != -999 | is.na(SPOSTMIN))
}
```

```{r}
rides_df <- rides[[1]]
for (i in 2:14) {
  rides_df <- rbind(rides_df, rides[[i]])
}
rides_df
```


The goal is to compare data between multiple dataframes so that only relevant data is used

Each ride has its own dataframe and is also part of a larger dataframe containing all rides

There is a metadata dataframe that has relevant data for each of the four parks

We want to create a new dataframe that has each ride as the row, the park its in, the ride opening data, if it has a splash aspect, and if it is indoors

***

> RSA: I am assuming the plot below is the one that needs to be reviewed.

Checking the dataframe being used:

```{r}
# checking for one of the parks
rides_df %>% 
  filter(ride_name == "dwarfs_train") %>% 
  mutate(
    myyear = year(datetime), 
    mymonth = month(datetime, label = TRUE), 
    myday = day(datetime), 
    myday_week = wday(datetime, label = TRUE), 
    myweek_month = week(datetime) # 
      
  ) %>% 
  group_by(myday_week, mymonth, myweek_month) %>% 
  summarise(mean_wait = mean(SPOSTMIN, na.rm = T))
```

Possible plot: 

```{r}
# checking for one of the parks
rides_df %>% 
  filter(ride_name == "dwarfs_train") %>% 
  mutate(
    myyear = year(datetime), 
    mymonth = month(datetime, label = TRUE), 
    myday = day(datetime), 
    myday_week = wday(datetime, label = TRUE), 
    myweek_month = week(datetime) # 
  ) %>% 
  group_by(myday_week, mymonth, myweek_month) %>% 
  summarise(mean_wait = mean(SPOSTMIN, na.rm = T)) %>% 
  ggplot() +
  geom_tile(color = "grey", aes(x = myday_week, y = myweek_month , fill = mean_wait)) +
  facet_wrap(~mymonth , scales = "free") +
  scale_fill_distiller(palette = "Spectral") +
  labs(fill = "Wait (min)", title = "Mean wait time by day of the year", 
       x = "", y = "") + 
  theme(legend.position = "bottom")
```

(you could think how to get the weeks of the month from the example above, so that instead of 1-52, it shows 1-5)

A calendar for one single year:

```{r}
rides_df %>% 
  filter(ride_name == "dwarfs_train") %>% 
  mutate(
    myyear = year(datetime), 
    mymonth = month(datetime, label = TRUE), 
    myday = day(datetime), 
    myday_week = wday(datetime, label = TRUE), 
    year_week = week(datetime)
    ) %>% 
  filter(myyear == 2019) %>% 
  group_by(mymonth, myday, myday_week, year_week) %>% 
  summarise(mean_wait = mean(SPOSTMIN, na.rm = T)) %>% 
  ggplot(aes(x = myday_week, y = year_week , fill = mean_wait)) +
  geom_tile(color = "grey") +
  geom_text(aes(label = myday), size = 2) + 
  facet_wrap(~mymonth , scales = "free") +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.text.y = element_blank()) +
  scale_fill_distiller(palette = "Spectral") +
  labs(fill = "Wait (min)", title = "Mean wait time by day for 2019", 
       x = "", y = "") + 
  theme(legend.position = "bottom")
```



To do: 
- Instead of weeks 1 - 52

***




```{r}
rides_df %>%
  mutate(year = year(as.POSIXct(mdy(date) + 1)), 
         month = month(as.POSIXct(mdy(date) + 1)), 
         day = weekdays(as.POSIXct(mdy(date) + 1)),
         monthweek = ceiling(day(mdy(date) + 1) / 7)) %>%
  group_by(month, day, monthweek) %>% 
  summarise(mean_wait = mean(SPOSTMIN, na.rm = TRUE)) 
```


```{r}
rides_df %>%
  mutate(year = year(as.POSIXct(mdy(date) + 1)), 
         month = month(as.POSIXct(mdy(date) + 1)), 
         day = weekdays(as.POSIXct(mdy(date) + 1)),
         monthweek = ceiling(day(mdy(date) + 1) / 7)) %>%
  group_by(month, day, monthweek) %>%
  summarise(mean_wait = mean(SPOSTMIN, na.rm = TRUE)) %>% 
  ggplot(aes(x = monthweek, y = ordered(day, levels = rev(days_of_week)), fill = mean_wait)) +
  geom_tile(color = "grey") +
  facet_grid(~month) +
  #scale_fill_gradient2(low = "blue", mid = "orange", high = "red", midpoint = 60) +
  scale_fill_distiller(palette = "Spectral") +
  labs(fill = "Wait (min)", title = "Mean wait time by day of the year for all rides", x = "Week of the month", y = "")
```

```{r, eval = FALSE}
rides_test <- rides_df

rides_test$weekday <- as.POSIXlt(mdy(rides_test$date))$wday

rides_test$weekdayf <- factor(rides_test$weekday,levels=rev(1:7),labels=rev(c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")),ordered=TRUE)

rides_test$monthf<-factor(month(mdy(rides_test$date)),levels=as.character(1:12),labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"),ordered=TRUE)

rides_test$yearmonth<- factor(as.yearmon(mdy(rides_test$date)))

rides_test$week <- as.numeric(format(mdy(rides_test$date),"%W"))

rides_test <- ddply(rides_test,.(yearmonth), transform, monthweek = 1 + week -min(week))

ggplot(rides_test, aes(monthweek, weekdayf, fill = SPOSTMIN)) + 
  geom_tile(colour = "white") + 
  facet_grid(~monthf) + 
  scale_fill_distiller(palette = "Spectral") + 
  xlab("Week of Month") + 
  ylab("") + 
  ggtitle("Time-Series Calendar Heatmap: AMZN Stock Prices") + 
  labs(fill = "Price") 
```

```{r, eval = FALSE}
rides_test <- rides_df
rides_test %>%
  mutate(
    weekday = as.POSIXlt(mdy(rides_test$date))$wday) %>%
  mutate(
    monthf = factor(month(mdy(rides_test$date)),levels=as.character(1:12),labels=c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"),ordered=TRUE),
    yearmonth = factor(as.yearmon(mdy(rides_test$date))),
    week = as.numeric(format(mdy(rides_test$date),"%W"))
  ) %>%
  ddply(.(yearmonth), transform, monthweek = 1 + week - min(week)) %>%
ggplot(aes(monthweek, factor(rides_test$weekday,levels=rev(1:7),labels=rev(c("Mon","Tue","Wed","Thu","Fri","Sat","Sun")),ordered=TRUE), fill = SPOSTMIN)) + 
  geom_tile(colour = "white") + 
  facet_grid(year(mdy(rides_test$date))~monthf) + 
  scale_fill_gradient(low="red", high="green") + 
  xlab("Week of Month") + 
  ylab("") + 
  ggtitle("Time-Series Calendar Heatmap: AMZN Stock Prices") + 
  labs(fill = "Price") 
```

